High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-18 notebooks/mnist.qmd
almost there

```{julia}
# Data:
counterfactual_data = load_mnist()
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
```

@@

```{julia}
epochs = 100
batchsize = Int(round(n_obs/10))

# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=MLJFlux.MLP(hidden=(32,), σ=relu), 
    epochs=epochs,
    batch_size=batchsize,
)

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=5)

# ResNet:
using MLJFlux, Metalhead
builder = MLJFlux.image_builder(ResNet, 18, pretrain=true)
resnet = ImageClassifier(
    builder=builder,
    epochs=epochs,
    batch_size=batchsize,
)

# ResNet Ensemble:
resnet_ens = EnsembleModel(model=resnet, n=5)
```

```{julia}

conf_model = conformal_model(mlp; method=:adaptive_inductive, coverage=.99)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

@@

# Generate counterfactual using generic generator:
generator = GenericGenerator()
ce_wachter = generate_counterfactual(
    x, target, dt_reduced, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=[0.0,10.0], 
    temp=0.01, 
    # opt=CounterfactualExplanations.Generators.JSMADescent(η=0.5),
)
ce_conformal = generate_counterfactual(
    x, target, dt_reduced, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

>>>>

```{julia}
# Data:
n_obs = 1000
counterfactual_data = load_mnist(n_obs)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# Model parameters:
epochs = 100
batch_size = Int(round(n_obs/10))
n_hidden = 32
activation = Flux.swish
builder = MLJFlux.MLP(hidden=(n_hidden,), σ=activation)
α = [0.33,1.0,1e-1]

# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=builder, 
    epochs=epochs,
    batch_size=batch_size,
)

# Joint Energy Model:
𝒟x = Uniform(-1,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=(input_dim,), batch_size=batch_size)
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    batch_size=batch_size,
    finaliser=Flux.softmax,
    loss=Flux.Losses.crossentropy,
    jem_training_params=(α=α,verbosity=10,),
    sampling_steps=20,
)

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=5)
```

```{julia}
cov = .9
conf_model = conformal_model(jem; method=:adaptive_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

```{julia}
jem = mach.model.model.jem
n_iter = 5000
_w = 1500
plts = []
neach = 10
for i in 1:10
    x = jem.sampler(jem.chain, jem.sampling_rule; niter=n_iter, n_samples=neach, y=i)
    plts_i = []
    for j in 1:size(x, 2)
        xj = x[:,j]
        xj = reshape(xj, (n_digits, n_digits))
        plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
    end
    plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
    plts = [plts..., plt]
end
Plots.plot(plts..., size=(_w,_w), layout=(10,1))
```

@@

# Generate counterfactual using generic generator:
generator = GenericGenerator()
ce_wachter = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=[0.0,10.0], 
    temp=0.01, 
    # opt=CounterfactualExplanations.Generators.JSMADescent(η=0.5),
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
